Untitled

### 2025-02-03 Drafts/Paper_1/02-intro.Rmd

Marine fishes were once believed to be essentially inexhaustible and thus extremely resilient to the effects of fishing [@mcintoshFisheriesInterNationalCouncil1919a]. However, high profile collapses of large fisheries and a subsequent lack of rebuilding (cite, cite) have fundamentally challenged this perception. The assumption of resilience originated from early theoretical population dynamics modeling [e.g., @verhulstNoticeLoiQue1838] that culminated in the development of modern fisheries science, largely built around the concept of maximum sustainable yield [MSY, @rickerStockRecruitment1954; @schaeferAspectsDynamicsPopulations1954; @bevertonDynamicsExploitedFish1957]. From its inception, MSY was controversial [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum197], yet the theoretical underpinning in equilibrium concepts and the associated idea that annual reproductive output and/or juvenile survival rates increase as abundance declines (implying harvesting activity will increase the per capita productivity of a population) persisted. The vast majority of modern assessments incorporate a stock-recruit function (cite, cite) making compensatory dynamics a foundational assumption about the life history of a species. Such fisheries assessments implicitly assume that lifetime reproductive rates in the absence of fishing would be expected to be greater than one (Myers cite, cite) and that even heavily-exploited fish populations retain a latent capacity for increase (cite, cite). Although MSY and equilibrium concepts are relatively simple and intuitive, and provide a means of developing tractable mechanistic models, questions remain regarding the applicability of these concepts and whether they adequately reflect the reality of the system being modeled [[@wormRebuildingGlobalFisheries2009; cite cite], particularly for species with markedly different life history characteristics (cite, cite).

Life history characteristics are now considered indicative of a species’ compensatory capacity (cite, cite) and are increasingly used within data-limited assessment approaches to provide management advice (cite). Theory suggests that small fishes, characterized by numerous offspring, rapid growth, early maturity and short lifespans, should have greater compensatory capacity, greater resilience to environmental variability and consequently, be better able to withstand fishing pressure [@kindsvaterTenPrinciplesEvolutionary2016; cite, cite]. Conversely, life histories characterized by longer lifespans and later maturity, or those with lower per-capita reproductive output are considered less resilient (cite). Furthermore, closely related species are anticipated to be more similar than distantly related species owing to a shared evolutionary history (Thorson fishlife), implying less heterogeneity among populations of a single species than among closely-related species. There is a general impression that forage fishes or populations at lower trophic levels are better able to take advantage of favorable environmental conditions and grow more quickly, particularly from low abundance (cite, cite). This resilience should manifest as a high latent capacity for population growth in the absence of fishing, as well as substantial variability in realized population growth rates among years. Dramatic increases in eulecan (FIND OTHER EXAMPLES) following years of particularly low abundance have provided empirical evidence of high compensatory capacity in forage fishes (cites). However, it remains unclear if correlated and compounding ecological and environmental processes (e.g. climate change effects, predatory prey dynamics) can largely obscure the importance of life history on realized productivity, making evolutionary history and life history characteristics a weak foundation for data-poor assessment in light of our changing oceans.

Stock assessment data from harvested marine telosts provide a unique opportunity to explore demographic variability and realized compensatory responses from a conservation biology or population ecology perspective. The long assessment history of these populations means both that their life history characteristics are well-known and that it is possible to quantify realized variability in survival and reproductive output, while accounting for the influence of fisheries. We based our analyses on a database containing age specific abundance information for 77 populations in the north Atlantic and the northeast Pacific Oceans [@charbonneauAgestructuredmarinefishdatabase2022]. By constructing life tables from abundance at age and catch at age data, we calculated common demographic metrics, including  annual population growth rates (cite), doubling time (cite), lifetime reproductive success (cite), and generation time (cite). Comparing the strength and magnitude of variability in realized productivity among species with different life histories as well as among populations of the same species yielded unexpected insights. Our results have implications for sustainable management of these exploited marine teleosts and could be used inform future management actions for populations that have collapsed. More broadly, our results challenged longstanding ideas on the applicability of equilibrium theory, life history theory, and compensatory dynamics to explain expected productivity in fishes.

>>>>

Marine fishes were once believed to be essentially inexhaustible and thus extremely resilient to the effects of fishing [@mcintoshFisheriesInterNationalCouncil1919a; @nielsenEvolutionFisheriesManagement1976]. However, high profile collapses of large fisheries and a subsequent lack of rebuilding [@hutchingsCollapseRecoveryMarine2000; @frankAlleeEffectsCompensatory2000; @myersMaximumPopulationGrowth1997] have fundamentally challenged this perception. The assumption of resilience originated from early theoretical population dynamics modeling [e.g., @verhulstNoticeLoiQue1838] that culminated in the development of modern fisheries science, largely built around the concept of maximum sustainable yield [MSY, @rickerStockRecruitment1954; @schaeferAspectsDynamicsPopulations1954; @bevertonDynamicsExploitedFish1957]. Despite long standing concerns regarding the applicability of MSY [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum1977], its theoretical underpinning in equilibrium concepts and the associated idea that annual reproductive output and/or juvenile survival rates increase as abundance declines persisted along with the implication that harvesting activity will increase the per capita productivity of a stock [@schaeferAspectsDynamicsPopulations1954]. Most modern assessments incorporate a stock-recruit function [@rickerStockRecruitment1954; @bevertonDynamicsExploitedFish1957; @hilbornQuantitativeFisheriesStock1992; @subbeyModellingForecastingStock2014] making compensatory dynamics a foundational assumption about the life history of a species. Such fisheries assessments implicitly assume that lifetime reproductive rates in the absence of fishing would be expected to be greater than one [@hilbornQuantitativeFisheriesStock1992; @myersPopulationDynamicsExploited1995] and that even heavily-exploited fish stocks retain a latent capacity for increase [maceRelationshipsCommonBiological1994]. Although MSY and equilibrium concepts are relatively straightforward, intuitive, and provide a means of developing tractable mechanistic models, questions remain regarding their applicability [@finleyAllFishSea2019; @larkinEpitaphConceptMaximum1977]. There are concerns about whether they accurately reflect the the system being modeled [@hilbornQuantitativeFisheriesStock1992; paulyPrimaryProductionRequired1995; @wormRebuildingGlobalFisheries2009], particularly for species with markedly different life history characteristics [hilbornFisheriesInducedChangesGrowth2008].

Life history characteristics are now considered indicative of a species’ compensatory capacity [@myersStockRecruitmentGeneralizations2001; @roseCompensatoryDensityDependence2001; @winemillerLifeHistoryStrategies2005] and are increasingly used within data-limited assessment approaches to provide management advice [@martellSimpleMethodEstimating2013; @pedersenStochasticSurplusProduction2017] . Theory suggests that small fishes, characterized by numerous offspring, rapid growth, early maturity and short lifespans, should have greater compensatory capacity, greater resilience to environmental variability and consequently, be better able to withstand fishing pressure [@kindsvaterTenPrinciplesEvolutionary2016; @lawFishingSelectionPhenotypic2000]. Conversely, life histories characterized by longer lifespans and later maturity, or those with lower per-capita reproductive output are considered less resilient [@winemillerLifeHistoryStrategies2005]. Furthermore, the life history traits of closely related species are anticipated to be more similar than those distantly related species owing to a shared evolutionary history [@thorsonPredictingRecruitmentDensity2020]. This implies there should be less heterogeneity in the productivity among stocks of a single species than among closely-related species. There is a general impression that forage fishes or stocks at lower trophic levels are better able to take advantage of favorable environmental conditions and grow more quickly, particularly from low abundance [@essingtonFishingMarineFood2006]. This resilience should manifest as a high latent capacity for population growth in the absence of fishing, as well as substantial variability in realized population growth rates among years. However, it remains unclear if correlated and compounding ecological and environmental processes (e.g. climate change effects, predator-prey dynamics) can largely obscure the importance of life history on realized productivity, making evolutionary history and life history characteristics a weak foundation for data-poor assessment in light of our changing oceans.

Stock assessment data from harvested marine telosts provide a unique opportunity to explore demographic variability and realized compensatory responses. The long assessment history of these stocks means that their life history characteristics are well-known and that it is possible to quantify realized variability in survival and reproductive output, while accounting the impact of fishery removals. We based our analyses on 77 marine fish stocks from a database of age-specific data [@charbonneauAgestructuredmarinefishdatabase2022]. By constructing life tables from abundance-at-age and catch-at-age data, we calculated common demographic metrics, including  annual population growth rates, doubling time, lifetime reproductive success, and generation time [@caswellMatrixPopulationModels2001]. Comparing the strength and magnitude of variability in realized productivity among species with different life histories as well as among stocks of the same species yielded unexpected insights. Our results have implications for sustainable management of these exploited marine teleosts and could be used to inform future management actions for stocks that have collapsed. More broadly, our results challenged longstanding ideas on the applicability of equilibrium theory, life history theory, and compensatory dynamics to explain expected productivity in fishes.

### 2025-02-03 Drafts/Paper_1/03-methods.Rmd

$$
RPA_{y,a} =  SSB_{(y-ra),a} RPS_{y}
$$
Fish are expected to have variability in reproductive success by age (ref, ref). However, we had to make the simplifying assumption that each kilogram of spawning biomass contributed the same to the total total reproductive output, irrespective of age. We calculated the age-specific fecundity ($f_{y,a}$) as the number of offspring produced by each spawner in the stock. The number of spawners in year $y$ and age class $a$ ($NS_{y,a]$) is the proportion of mature individuals in each age class ($Mat_{y,a}$) multiplied by the number of individuals in each age class:

>>>>

$$
RPA_{y,a} =  SSB_{(y-ra),a} RPS_{y}
$$
Fish are expected to have variability in reproductive success by age (ref, ref). However, we had to make the simplifying assumption that each kilogram of spawning biomass contributed the same to the total total reproductive output, irrespective of age. We calculated the age-specific fecundity ($f_{y,a}$) as the number of offspring produced by each spawner in the stock. The number of spawners in year $y$ and age class $a$ ($NS_{y,a}$) is the proportion of mature individuals in each age class ($Mat_{y,a}$) multiplied by the number of individuals in each age class:

### 2025-02-03 Drafts/Paper_1/04-results.Rmd

The overall median population growth rate when excluding the effects of fishing ($\lambda_{F=0}$) was `r med.lambda.F0`, which represents a latent capacity for population growth of only `r med.r.F0`% annually. Aggregating over years, the population with highest median theoretical population growth rate was Atlantic cod in the Irish Sea (median $\lambda_{F=0}$ = `r max.median.lam.nf`) and the lowest was pollock in the Gulf of Alaska (median $\lambda_{F=0}$ = `r min.median.lam.nf`).  When the effects of fishing was were accounted for, the overall median population growth rate ($\lambda_{F>0}$) declined to `r med.lambda.F`, which would correspond to overfishing by  `r -med.r.F`% annually. The realized range of population growth rates was much narrower, as would be expected from managed stocks, where a typical goal is to maintain stock size at or above a specific biomass reference point [CITE]. Indeed, median $\lambda_{F>0}$ was below one the majority of the time for `r prop.usually.lt.1.fish`% of the stocks, suggesting that stocks were being managed in a way that essentially eliminated their capacity for abundance increase in more than half of the years in the data. However, many of the populations included in our analysis experienced infrequent large recruitment events that resulted in high $\lambda_{F>0}$ estimates (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-bx-other-fig)). For example, while Gulf of Alaskan pollock has the lowest median $\lambda_{F=0}$, it's mean $\lambda_{F=0}$ was `r mean.nf.of.min.median.lam.nf`, and when the effects of fishing are included this population had the $8^{th}$ highest mean $\lambda_{F>0}$ (`r mean.fish.of.min.median.lam.nf`) in this analysis. Consequently, the overall mean realized growth rate $\lambda_{F>0}$ for the populations in this analysis remained slightly greater than one. On average, population growth tended to compensate for declines in other years, maintaining a net positive overall productivity for most of these stocks.

@@

Overall, the data suggested that negative density dependence was common as both $\lambda_{F=0}$ and $\lambda_{F>0}$ increased as abundance declined (Figure \@ref(fig:dd-overall-plt)). Changes in the rate of population growth were more pronounced at low and high abundance levels; the relationship from the fitted GAMM was essentially flat when relative abundance ranged between 40-70% of maximum. This suggests that there was a surprisingly large range of population abundances that would not be expected to exhibit strong compensatory dynamics. The predicted population growth rate excluding the impacts of fishing ($\lambda_{F=0}$) only declined below one when the populations approached their maximum observed abundance ($\gtrsim$ `r pa.lam.nf.at.one`%). However, when accounting for fishing mortality ($\lambda_{F>0}$), the modelled relationship dropped below one when the population abundance exceeded approximately `r pa.lam.fish.at.one`% of their maximum abundance. At lower abundances, the model estimated population growth rate excluding fishing ($\lambda_{F=0}$) increased more slowly than realized population growth rates ($\lambda_{F>0}$) likely due to reductions in fishing mortality at low abundance. However, the fitted values never converged, suggesting that some fishing mortality remained even at very low abundance (Figure \@ref(fig:dd-overall-plt)).

As expected, there was generally an increase in the variability of the $\lambda$ estimates as abundance declines, largely driven by an increase in the magnitude of the maximum value of $\lambda$ and a small increase in the frequency of low $\lambda$ values (Figure \@ref(fig:dd-overall-plt)). While there was a consistent overall pattern, the estimated random effects demonstrated substantial variability among stocks in the strength of negative density dependence. For examples, populations in the western Atlantic (particularly Atlantic cod) and Alaska had some of the lowest $\lambda_{F=0}$ estimates at low abundance, while several populations in the North Sea (including Atlatnic cod) had among the highest $\lambda_{F=0}$ estimates at low abundance (Table \@ref(tab:tab-lambdas)).  Of the `r n.bl.40` populations that had declined below 40% of the maximum abundance, only `r n.low.40.nf` of the populations had $\lambda_{F=0}$ estimates < 1 at 40% of their maximum abundance; when the effects of fishing were included `r n.low.40.nf` of the populations had $\lambda_{F>0}$ estimates below 1 (Table \@ref(tab:tab-lambdas)).

The vast majority of populations the lowest values of both $\lambda_{F=0}$ (`r n.low.lam.at.max.nf` of `r n.stocks`) and $\lambda_{F>0}$ (`r n.low.lam.at.max.fish` of `r n.stocks`) were found near the maximum abundance in the time series, this conforms to the expectations of this type of historical reconstruction and population dynamics theory (i.e., negative density dependence; see Supplemental Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt)). However, Atlantic cod in the Northwest Atlantic comprised `r n.cod.lam.not.at.max.nf` of the `r n.stocks - n.low.lam.at.max.nf` populations in which the lowest population growth rates were observed at abundances markedly below the maximum. These Atlantic cod populations seemed to have the least compensatory capacity at relative abundances between `r pa.low.cod.min` and `r pa.low.cod.max` even when excluding the impacts of fishing mortality. For two of the `r n.stocks - n.low.lam.at.max.nf` populations, the lowest $\lambda_{F=0}$ estimates were less than 0.005 smaller than the estimate at the maximum abundance, indicating extremely similar compensatory capacity over the majority of observed stock sizes. The declines in $\lambda_{F=0}$ for the other four populations were more pronounced (Supplemental Figure \@ref(fig:dd-nf-stock-plt)), ranging from `r low.4.lam.40.nf` to `r low.lam.40.nf` (Tables \@ref(tab:tab-lambdas) and \@ref(tab:tab-lamda-diffs)).

@@

Lifetime reproductive output and generation length were estimated for each cohort in which data were available, population decline would be assured from lifetime reproductive rates consistently below 1. When excluding fishing mortality, the overall median lifetime reproductive output, (i.e., the number of females spawners per female spawner) was `r r0.med` (mean = `r r0.mn`; Figure \@ref(fig:r0-6-plt)). When realized fishing mortality was accounted for, the median reproductive output dropped to `r r0.med.fish` (mean = `r r0.mn.fish`). This difference between the mean and median for stocks experiencing exploitation, again highlighted the importance of large reproductive events to persistence by maintaining the latent capacity for population growth (Figure \@ref(fig:r0-6-plt)). For species which were found across the North Atlantic, the lifetime reproductive output when excluding fishing mortality tended to decline from east to west, which again was most evident for the Atlantic cod populations. While the reproductive potential of cohorts within a population could be highly variable, there were no clear temporal trends in $R_0$ (Figures \@ref(fig:r0-nf-ts-plt) and \@ref(fig:r0-fish-ts-plt)).

Generation length (calculated as the mean age of spawners contributing to a recruit cohort) tended to become more similar with fishing pressure included both among stocks and among species (Figures \@ref(fig:gl-6-plt) and \@ref(fig:gl-other-plt)). This is an expected outcome of reduced survivorship-to-age. However, there were regional and interspecific patterns that were unexpected. For example, Atlantic herring is typically considered to be a short-lived species yet its generation length in the Central and western Atlantic tended to be higher than the generation length of Atlantic cod and haddock (especially for the populations in the eastern Atlantic; (Figure \@ref(fig:gl-6-plt)). Similar to lifetime reproductive output there were no obvious temporal trends in the generation length among cohorts for the populations in this analysis  (Figures \@ref(fig:gl-nf-ts-plt) and \@ref(fig:gl-fish-ts-plt)).

Based on simulated stock trajectories, the recovery potential in the absence of fishing was relatively high for the majority of species (Figure \@ref(fig:dub-time-map)). Of the `r n.stocks` populations in this analysis, `r n.dub.10` had at least 75% of the simulations double in size withing 10 years, which increased to `r n.dub.20` populations by 20 years (Figure \@ref(fig:dub-ts)). Conversely, there were clear regional trends in stocks with low recovery potential. Of the `r n.not.dub.20` populations in this analysis in which < 50% of simulations doubled in size after 20 years, `r n.alaska.slow` were in Alaska and `r n.nwa.slow` in NAFO Regions 2-4 in the western Atlantic. This meant that `r prop.slow.alaska`% and `r prop.slow.n24`% of assessed stocks in Alaska and NAFO Regions 2-4, respectively, had low recovery potential. This contrasts to only `r n.nea.slow` of the `r n.nea` populations in the eastern Atlantic with a low recovery potential within 20 years in the absence of fishing (Figure \@ref(fig:dub-time-map)).

Recovery potential was intrinsically linked to both the magnitude and variability in annual population growth rates. The recovery potential was influenced both b$\lambda_{F=0}$ and the variability in $\lambda_{F=0}$. Populations with a median $\lambda_{F=0} < 1$  had moderate (defined as populations in which 50-75% of the simulations doubled within 20 years) to high (> 75% doubled in size within 20 years) recovery potential whenever the standard deviation in $\lambda$ was $\gtrsim$ 0.4 (Figure \@ref(fig:dub-vs-sd)). Populations with a median $\lambda_{F=0}$ near 1 and with low variability tended to be stocks with low recovery potential.

>>>>

The overall median population growth rate when excluding the effects of fishing ($\lambda_{F=0}$) was `r med.lambda.F0`, which represents a latent capacity for population growth of only `r med.r.F0`% annually. Aggregating over years, the population with highest median theoretical population growth rate was Atlantic cod in the Irish Sea (median $\lambda_{F=0}$ = `r max.median.lam.nf`) and the lowest was pollock in the Gulf of Alaska (median $\lambda_{F=0}$ = `r min.median.lam.nf`).  When the effects of fishing was were accounted for, the overall median population growth rate ($\lambda_{F>0}$) declined to `r med.lambda.F`, which would correspond to overfishing by  `r -med.r.F`% annually. The realized range of population growth rates was much narrower, as would be expected from managed stocks, where a typical goal is to maintain stock size at or above a specific biomass reference point [@marentetteCrossJurisdictionalReviewInternational2020]. Indeed, median $\lambda_{F>0}$ was below one the majority of the time for `r prop.usually.lt.1.fish`% of the stocks, suggesting that stocks were being managed in a way that essentially eliminated their capacity for abundance increase in more than half of the years in the data. However, many of the populations included in our analysis experienced infrequent large recruitment events that resulted in high $\lambda_{F>0}$ estimates (Figures \@ref(fig:lambda-bx-6-fig) and \@ref(fig:lambda-bx-other-fig)). For example, while Gulf of Alaskan pollock has the lowest median $\lambda_{F=0}$, it's mean $\lambda_{F=0}$ was `r mean.nf.of.min.median.lam.nf`, and when the effects of fishing are included this population had the $8^{th}$ highest mean $\lambda_{F>0}$ (`r mean.fish.of.min.median.lam.nf`) in this analysis. Consequently, the overall mean realized growth rate $\lambda_{F>0}$ for the populations in this analysis remained slightly greater than one. On average, population growth tended to compensate for declines in other years, maintaining a net positive overall productivity for most of these stocks.

@@

Overall, the data suggested that negative density dependence was common as both $\lambda_{F=0}$ and $\lambda_{F>0}$ increased as abundance declined (Figure \@ref(fig:dd-overall-plt)). Changes in the rate of population growth were more pronounced at low and high abundance levels; the relationship from the fitted GAMM was essentially flat when relative abundance ranged between 40-70% of maximum. This suggests that there was a surprisingly large range of stock abundances that would not be expected to exhibit strong compensatory dynamics. The predicted population growth rate excluding the impacts of fishing ($\lambda_{F=0}$) only declined below one when the stocks approached their maximum observed abundance ($\gtrsim$ `r pa.lam.nf.at.one`%). However, when accounting for fishing mortality ($\lambda_{F>0}$), the modelled relationship dropped below one when the stock abundance exceeded approximately `r pa.lam.fish.at.one`% of their maximum abundance. At lower abundances, the model estimated population growth rate excluding fishing ($\lambda_{F=0}$) increased more slowly than realized population growth rates ($\lambda_{F>0}$) likely due to reductions in fishing mortality at low abundance. However, the fitted values never converged, suggesting that some fishing mortality remained even at very low abundance (Figure \@ref(fig:dd-overall-plt)).

As expected, there was generally an increase in the variability of the $\lambda$ estimates as abundance declines, largely driven by an increase in the magnitude of the maximum value of $\lambda$ and a small increase in the frequency of low $\lambda$ values (Figure \@ref(fig:dd-overall-plt)). While there was a consistent overall pattern, the estimated random effects demonstrated substantial variability among stocks in the strength of negative density dependence. For examples, stocks in the western Atlantic (particularly Atlantic cod) and Alaska had some of the lowest $\lambda_{F=0}$ estimates at low abundance, while several stocks in the North Sea (including Atlatnic cod) had among the highest $\lambda_{F=0}$ estimates at low abundance (Table \@ref(tab:tab-lambdas)).  Of the `r n.bl.40` stocks that had declined below 40% of the maximum abundance, only `r n.low.40.nf` of the stocks had $\lambda_{F=0}$ estimates < 1 at 40% of their maximum abundance; when the effects of fishing were included `r n.low.40.nf` of the stocks had $\lambda_{F>0}$ estimates below 1 (Table \@ref(tab:tab-lambdas)).

The vast majority of stocks the lowest values of both $\lambda_{F=0}$ (`r n.low.lam.at.max.nf` of `r n.stocks`) and $\lambda_{F>0}$ (`r n.low.lam.at.max.fish` of `r n.stocks`) were found near the maximum abundance in the time series, this conforms to the expectations of this type of historical reconstruction and population dynamics theory (i.e., negative density dependence; see Supplemental Figures \@ref(fig:dd-nf-stock-plt) and \@ref(fig:dd-fish-stock-plt)). However, Atlantic cod in the Northwest Atlantic comprised `r n.cod.lam.not.at.max.nf` of the `r n.stocks - n.low.lam.at.max.nf` stocks in which the lowest population growth rates were observed at abundances markedly below the maximum. These Atlantic cod stocks seemed to have the least compensatory capacity at relative abundances between `r pa.low.cod.min` and `r pa.low.cod.max` even when excluding the impacts of fishing mortality. For two of the `r n.stocks - n.low.lam.at.max.nf` stocks, the lowest $\lambda_{F=0}$ estimates were less than 0.005 smaller than the estimate at the maximum abundance, indicating extremely similar compensatory capacity over the majority of observed stock sizes. The declines in $\lambda_{F=0}$ for the other four stocks were more pronounced (Supplemental Figure \@ref(fig:dd-nf-stock-plt)), ranging from `r low.4.lam.40.nf` to `r low.lam.40.nf` (Tables \@ref(tab:tab-lambdas) and \@ref(tab:tab-lamda-diffs)).

@@

Lifetime reproductive output and generation length were estimated for each cohort in which data were available, population decline would be assured from lifetime reproductive rates consistently below 1. When excluding fishing mortality, the overall median lifetime reproductive output, (i.e., the number of females spawners per female spawner) was `r r0.med` (mean = `r r0.mn`; Figure \@ref(fig:r0-6-plt)). When realized fishing mortality was accounted for, the median reproductive output dropped to `r r0.med.fish` (mean = `r r0.mn.fish`). This difference between the mean and median for stocks experiencing exploitation, again highlighted the importance of large reproductive events to persistence by maintaining the latent capacity for population growth (Figure \@ref(fig:r0-6-plt)). For species which were found across the North Atlantic, the lifetime reproductive output when excluding fishing mortality tended to decline from east to west, which again was most evident for the Atlantic cod stocks. While the reproductive potential of cohorts within a stock could be highly variable, there were no clear temporal trends in $R_0$ (Figures \@ref(fig:r0-nf-ts-plt) and \@ref(fig:r0-fish-ts-plt)).

Generation length (calculated as the mean age of spawners contributing to a recruit cohort) tended to become more similar with fishing pressure included both among stocks and among species (Figures \@ref(fig:gl-6-plt) and \@ref(fig:gl-other-plt)). This is an expected outcome of reduced survivorship-to-age. However, there were regional and interspecific patterns that were unexpected. For example, Atlantic herring is typically considered to be a short-lived species yet its generation length in the Central and western Atlantic tended to be higher than the generation length of Atlantic cod and haddock (especially for the stocks in the eastern Atlantic; (Figure \@ref(fig:gl-6-plt)). Similar to lifetime reproductive output there were no obvious temporal trends in the generation length among cohorts for the stocks in this analysis  (Figures \@ref(fig:gl-nf-ts-plt) and \@ref(fig:gl-fish-ts-plt)).

Based on simulated stock trajectories, the recovery potential in the absence of fishing was relatively high for the majority of species (Figure \@ref(fig:dub-time-map)). Of the `r n.stocks` stocks in this analysis, `r n.dub.10` had at least 75% of the simulations double in size withing 10 years, which increased to `r n.dub.20` stocks by 20 years (Figure \@ref(fig:dub-ts)). Conversely, there were clear regional trends in stocks with low recovery potential. Of the `r n.not.dub.20` stocks in this analysis in which < 50% of simulations doubled in size after 20 years, `r n.alaska.slow` were in Alaska and `r n.nwa.slow` in NAFO Regions 2-4 in the western Atlantic. This meant that `r prop.slow.alaska`% and `r prop.slow.n24`% of assessed stocks in Alaska and NAFO Regions 2-4, respectively, had low recovery potential. This contrasts to only `r n.nea.slow` of the `r n.nea` stocks in the eastern Atlantic with a low recovery potential within 20 years in the absence of fishing (Figure \@ref(fig:dub-time-map)).

Recovery potential was intrinsically linked to both the magnitude and variability in annual population growth rates. The recovery potential was influenced both b$\lambda_{F=0}$ and the variability in $\lambda_{F=0}$. Stocks with a median $\lambda_{F=0} < 1$  had moderate (defined as stocks in which 50-75% of the simulations doubled within 20 years) to high (> 75% doubled in size within 20 years) recovery potential whenever the median $\lambda_{F=0}$ was $\gtrsim$ `r ` lam.rebuild.safe or the standard deviation in $\lambda$ was $\gtrsim$ `r sd.rebuild.safe` (Figure \@ref(fig:dub-vs-sd)). Stocks with a median $\lambda_{F=0}$ near 1 and with low variability tended to be stocks with low recovery potential.

### 2025-02-03 Drafts/Paper_1/07-tables.Rmd

>>>>

```{r, process}
# This will be the meta data really...
# remove columns we don't need.

rm.meta <- which(names(meta.sub) %in% c("Meeting_or_reference","G.short","G.Species","Manage.short","LME","Area.broad","Area.long","Area.narrow","Stock.short","Stock","Model","Case","Notes","Area"))
meta.4.tab <- meta.sub[,-rm.meta]
re.ord <- c(8,2:7,1)
meta.4.tab <- meta.4.tab[,re.ord]
names(meta.4.tab) <- c("Stock", "Order", "Family", "Genus", "Species", "Minimum Age", "Maximum Age", "Management Body")

```

```{r,tab-lambdas}
# The table with the lambda values
csas_table(comp.lam.res.tab,caption ="A comparison of the $\\lambda_{F=0}$ and $\\lambda_{F>0}$ estimates for stock at the minimum and maximum abundance, along with the estimate at 40\\% of the maximum abundance. values greater than 1 indicate that the population would be expected to increase in size, while values less than 1 indicate the population is likely to decline in size")
```

\newpage

```{r,tab-lamda-diffs}
# Now make the table for the differences
csas_table(comp.lam.diff.res.tab,caption ="A comparison of the difference in $\\lambda_{F=0}$ and $\\lambda_{F>0}$ for each stock at the minimum and maximum abundance, along with the change at 40\\% of the maximum abundance. Positive values indicate that the $\\lambda$ estimae has increased as abundance declines (consistent with negative density dependence), while negative values indicate the $\\lambda$ estimate has declined (considtent with positive density dependence)")
```

```{r,tab-meta}
# Now make the table for the differences
csas_table(meta.4.tab,caption ="Summary of the stocks included in this analysis.  The minimun and maximum ages are what was used in the stock assessment.")
```

### 2025-02-03 Drafts/Paper_1/08-figures.Rmd

\newpage
```{r lambda-bx-other-fig, echo=F, fig.cap="$\\lambda$ estimates for the remainder of the stocks in this analysis, plotted by the location of the stock. The abreviated species name is provided on the x axis."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_other_species_",what.2.tune,".png"))

>>>>

\newpage
```{r lambda-bx-other-fig, echo=F, fig.cap="$\\lambda$ estimates for the remainder of the stocks in this analysis, plotted by the location of the stock. The abbreviated species name is provided on the x axis."}
knitr::include_graphics(paste0(loc,"/Figures/tuned/No_age_correction/Lambda_boxplot_other_species_",what.2.tune,".png"))

### 2025-02-03 Scripts/Test_for_danielle.Rmd
more more more

  So now I can try to do some tuning, both forwards and backwards, lets see what we get from that...

```{r tunes,include=F,echo=F,warning=F,message=F}

@@

     for.tune.all.res[[i]] <- fast.for.tune
     for.tune.tmp <- data.frame(fast.for.tune$res,Stock = i)
     for.tune.summary[[i]] <- for.tune.tmp
  #}

}# end if load results

>>>>

@@

     for.tune.all.res[[i]] <- fast.for.tune
     for.tune.tmp <- data.frame(fast.for.tune$res,Stock = i)
     for.tune.summary[[i]] <- for.tune.tmp
  #}
```
